\hspace{1cm} Do syntactic and semantic similarity lead to interference effects? \newline Evidence from self-paced reading and event-related potentials using German

### 2024-08-24 paper/ResponseToReviewers.tex

\newpage

\subsection*{Comment R1.5}
\begin{quote}
``Another problem I have is more general and has to do with progress in science. The authors in my reading find the current findings more convincing than, say, Mertzen et al. study. I get that the current study is convincing (I do find it convincing) but that does not disqualify what Mertzen et al. found in any way - both results co-exist. Ideally, I would expect that a model would be able to explain the current data, but also the past data. I get it that this is not always possible but then, \textbf{it would be good if the authors make their position explicit - either by admitting that only some data can be captured, or, if the authors believe that the past data should be ignored, it should be explained why this is so.}

All in all, I think the interpretation of the data and the model have to be strengthened to meet requirements 1-3. Alternatively, the manuscript should explicitly admit issues in the proposed interpretation - that is, explain that the model fails to capture encoding interference (point 1 if I am right), explain why the model underpredicts interference in syntax/interaction and explain why it is enough, at least for now, to only model the current findings and to put previous findings aside (assuming they actually differ - see point 3 about Bayes factors above).''\end{quote}

\subsubsection*{Response to R1.5}
About why our results diverge from previous studies' and whether we think our results are correct and the previous work is not, we have no position on this. We are simply reporting what we found. Contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{lcpaape2023, taboretal04}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree}, agreement attraction in grammatical sentences \citep{wagers2009agrttr, nicenboim} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}. The only way to resolve contradictions in the literature is by carrying out adequately powered studies in the future that can distinguish between the theoretical claims. Our work does not stand out as an anomaly; if anything, it's quite the opposite. We have added a discussion of divergent results in the literature to page \pageref{not_an_anomaly} in the General Discussion:

@@

\subsection*{Comment R1.8}
\begin{quote}
``Priors for Bayes factors\\
=============

Various priors are considered for Bayes factors. \textbf{Of course, it is good to explore several values but I wonder whether it is justified to conclude that syntactic interference or interaction should be excluded if they are not supported in models with unreasonably large priors (e.g., N+(0, 0.1)). Would anybody expect that interference is in range of 0.1 log-ms? As one of the authors of this manuscript nicely showed in Jager et al. (2017) and Schad et al. (2022), interference effects are small (at least for subject-verb agreement and ungrammatical illusions). Given that, I don't understand why we require a BF in favor of interference for a prior that is of size N+(0, 0.1). Shouldn't a smaller prior distribution suffice as evidence? This brings me again to my confusion in what the experiments actually show. The authors seem to argue that the only interference is the semantic one and they base the conclusion on large priors. But focusing on small priors (N+(0, 0.05) for spr, N-(0, 0.5) for EEG)), one could conclude that the interaction/syntactic interference might be present after all. Is there an independent reason to select the former priors over the latter for one's interpretation? To be sure, I don't want the authors to change the interpretation, but I would like to see either a more explicit argumentation why we should base our conclusion on such and such priors, or a more reflection in the text that explicitly states that the results are not equivocal and the current study is still compatible with a position that syntactic interference and/or interaction are real.}''\end{quote}

\subsubsection*{Response to R1.8}
We agree that syntactic interference and/or the interaction could be real, but our data does not provide conclusive evidence for them. Because it can be difficult to discern small differences in the Bayes factor plots (Figures 5 and 11 in Response to E.2), the Table below shows all Bayes factors for syntactic interference and the interaction. All Bayes factors above 1 are highlighted in bold. Following \textcite{jeffreys1998theory} and \textcite{lee2014bayesian}, we regard Bayes factors around 1 as providing no evidence. The only Bayes factor remarkably above 1 is BF$_{10}$ = 2.8 for the syntactic interference effect in the reading times of the spill-over region. Because of the pre-critical effects in the reading times, this effect in the spill-over region is confounded. Therefore, we think that although the Bayes factor theoretically provided anecdotal evidence for syntactic interference in the spill-over region, there is no \textit{decisive} evidence for syntactic interference in our data.

\begin{table}[h]
    \centering
    \begin{tabular}{ccccc}
\toprule
Experiment &      Region   &   Effect  &         Prior & BF10\\
\midrule
SPR & pre-critical&   syntactic &Normal(0, 0.01)&0.5\\
SPR & pre-critical&   syntactic &Normal(0, 0.05)& 0.11\\
SPR & pre-critical&   syntactic  &Normal(0, 0.1)& 0.06\\
SPR & critical   &syntactic &Normal(0, 0.01)&0.5\\
SPR & critical   &syntactic &Normal(0, 0.05)& 0.11\\
SPR & critical   &syntactic  &Normal(0, 0.1)&0.05\\
SPR & spill-over  & syntactic &Normal(0, 0.01)& \textbf{2.8}\\
SPR & spill-over  & syntactic &Normal(0, 0.05)& 0.8\\
SPR & spill-over  & syntactic  &Normal(0, 0.1)&0.42\\

EEG & N400   &syntactic &Normal(0, 0.1)&\textbf{1.18}\\
EEG & N400   &syntactic &Normal(0, 0.5)&0.52\\
EEG &  N400  & syntactic &  Normal(0, 1)&0.27\\
EEG & N400   &syntactic  & Normal(0, 2)&0.14\\
EEG & P600  & syntactic &Normal(0, 0.1) & 0.91\\
EEG & P600  & syntactic &Normal(0, 0.5) &0.27\\
EEG &  P600 &  syntactic &  Normal(0, 1)&0.14\\
EEG & P600  & syntactic  & Normal(0, 2) & 0.07\\

SPR & pre-critical &interaction &Normal(0, 0.01) & 1\\
SPR & pre-critical &interaction &Normal(0, 0.05) & 0.36\\
SPR & pre-critical &interaction & Normal(0, 0.1) & 0.18\\
SPR & critical &interaction &Normal(0, 0.01) & \textbf{1.36}\\
SPR & critical &interaction &Normal(0, 0.05) & 0.71\\
SPR & critical &interaction & Normal(0, 0.1) & 0.37\\
SPR & spill-over& interaction& Normal(0, 0.01) &0.92\\
SPR & spill-over &interaction &Normal(0, 0.05) & 0.3\\
SPR & spill-over &interaction  &Normal(0, 0.1) & 0.15\\

EEG & N400 &interaction &Normal(0, 0.1)&0.88\\
EEG & N400 &interaction &Normal(0, 0.5)&0.36\\
EEG & N400 &interaction  & Normal(0, 1)&0.19\\
EEG & N400 &interaction  & Normal(0, 2) & 0.1\\
EEG & P600 &interaction &Normal(0, 0.1)& \textbf{1.16}\\
EEG & P600 &interaction &  Normal(0, 1) &0.43\\
EEG & P600 &interaction &  Normal(0, 2) &0.22\\
\bottomrule
\end{tabular}

\end{table}

%BFs do not to need to be strongly in favor of an effect across all investigated priors to conclude that an effect is supported. However, to draw decisive conclusions in favor of an effect, the evidence for that effect should be stronger than anecdotal for at least a subset of reasonable priors.

Regarding the choice of priors, we believe that it is reasonable to explore a wide range of a priori effect sizes, specifically from $\pm$8 to $\pm$80 ms for the reading times. The meta-analysis of reading time studies by \textcite{jaeger_etal_2017} showed that the pooled estimate of inhibitory interference across studies with similar designs as ours was indeed small (13.1 ms, page 327), but some studies showed effects around 50 ms (e.g., two of the sub-experiments from \cite{vandyke07}) and one experiment by Franck et al., (2015) even showed an estimate above 100 ms (see Figure 2 on page 329 of \cite{jaeger_etal_2017}). Because such large interference effects have been found previously (see also our new Figure \ref{fig:previous_vs_present} showing all estimates of \cite{vandyke07} and \cite{mertzen}).

@@

\subsubsection*{Response to R3.2}
The Reviewer is correct, our proposal that hierarchical syntactic structure is used during retrieval relies on the general assumption that hierarchical syntactic structure is being tracked. We think that this can easily be represented in encodings and then be used during retrieval. We have added a more explicit description of this underlying mechanism to page \pageref{clause_tracker}:

\begin{quote}
    \Paste{clause_tracker}
\end{quote}

\subsection*{Comment R3.3}
\begin{quote}
``\textbf{I also have some questions or reservations about the account of semantic interference effects in terms of features like [+animate]. This is a straightforward enough feature to invoke, and it is a plausible content feature for noun phrases. But I am less sure how well it generalizes. There are many situations in which verb-noun combinations are intuitively a good fit, but where it is less easily cast in terms of features and retrieval cues. For example, we share the intuition that ``the teacher drove …" is a better combination than ``the little boy drove …", because teachers are plausible drivers, whereas small children are not. This is less easily captured in terms of features and retrieval cues. Maybe the verb ``drive" wants a subject with a [+driver] feature. But is this plausibly something that is part of the memory encoding of "teacher" but not ``little boy"? Does this require something more than a toy theory of semantic encoding? Or does it predict that only simple features like [+animate] are able to cause semantic interference?}''
\end{quote}

\subsubsection*{Response to R3.3}
It has been shown in the literature that fine-grained semantic features, such as [+fixable] \citep{vandyke_mcelree06} and [+shatterable] \citep{cunnings_sturt_2018retrieval}, are used during cue-based retrieval. The materials in the design employed here were constructed so that the potentially fine-grained features relevant for the situation at hand (e.g., [+can drive], [+can drink] or [+can smile]) could be summed under the umbrella feature [+animate] or actually [+human]. To resolve the critical dependency it was always sufficient to use the umbrella feature [+animate] (or [+human]). The subject and the animate distractor always shared the [+animate] (or more precisely [+human]) feature, while the inanimate distractor was always [-animate] and hence [-human]. There were no animate distractors which mismatched a fine-grained verb-specific retrieval cue like [+can drive]. The results of the plausibility norming, see Figure \ref{fig:plausibility_rating} on page \pageref{fig:plausibility_rating} of the manuscript illustrate that whatever semantic feature determines a good fit between subject and verb matched with the features of the subject and the animate distractor, but not with the features of the inanimate distractor. For a principled approach to feature selection within the cue-based retrieval framework, see \textcite{smith2020principled}.

### 2024-08-24 paper/main.tex
Updated via the Open Science Framework

Semantic interference is manipulated via the animacy of the distractor. During the retrieval of the subject of the verb \textit{trank} (`drank'), the animate distractor  \textit{Einbrecher} (`burglar') is assumed to cause high semantic interference, while the inanimate distractor \textit{Verlust} (`loss') is assumed to cause low semantic interference. \Copy{semantic_features_encoding}{\textcolor{blue}{The animacy of the distractor is used as a stand-in feature for semantic features relevant for the specific retrieval event which might be more fine-grained. In \ref{ex:materials}, the semantic retrieval cue might be \{+animate\} or \{+human\} but it might as well be \{+can drink\}. While more generic features like \{+animate\} or \{+human\} (all animate distractors were human) might be created during memory encoding, more specific features like \{+can drink\} are probably only activated when relevant. The present study is not designed to provide insight into the specific semantic features used during memory encoding and retrieval (for work on highly specific retrieval cues like \{+sailable\} and \{+shatterable\}, see \cite{vandyke_mcelree06, cunnings_sturt_2018retrieval}). A plausibility norming experiment (see page \pageref{plausib_normin_orig}) was conducted to ensure that the semantic manipulation did work, i.e., that the semantic requirements of the verb matched the features of the subject and the animate distractor, but not the features of the inanimate distractor.
}}\label{semantic_features_encoding}

@@

Was there any evidence for the effects \textcolor{blue}{of interest }estimated above? The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}. The analyses use all the data from Experiments 1a and 1b. \Copy{SPR_bfs}{\textcolor{blue}{We found very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231). In the critical region, the evidence for semantic interference was inconclusive across the range of investigated priors (BF$_{10}$ between 0.7 and 3.5). There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5). Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8). Similarly,} \Copy{insig_interaction}{\textcolor{blue}{there was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1). Only under the narrowest prior in the critical region, there was very weak evidence for the interaction (BF$_{10}$ = 1.4).}}\label{insig_interaction}

\textcolor{blue}{All in all, the Bayes factor analysis provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but inconclusive results in the critical region. In contrast, the Bayes factor analysis provided at best weak evidence for syntactic interference and the interaction or even evidence against them.}

\Copy{fig_bf_spr}{\begin{figure}[htpb]
    \caption{Bayes Factors for the effects of semantic interference, syntactic interference and their interaction in the reading times (combined Experiments 1a and 1b) at the critical verb and surrounding regions, under a range of priors on the target parameters.}
    \label{fig:spr_bfs}
    \centering
    \includegraphics[width=\textwidth]{images/BF_plot_spr_774_allregions.png}
\end{figure}
}}

\subsection{Discussion}
We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation. \textcolor{blue}{Reading times of the pre-critical word and critical verb showed evidence against syntactic interference. In the spill-over region, there was weak evidence for a small syntactic interference effect in the size of up to $\pm$8 ms. We will discuss the almost complete absence of syntactic interference in our data in the general discussion. The reading time results at the critical verb were inconclusive for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region. This means that semantic interference occurred even before the retrieval event could begin. Reading times at the critical verb showed very weak support for a small interaction of syntactic and semantic interference. There was evidence against the interaction under wider priors and in the surrounding regions.} \Copy{only_precritical}{\textcolor{blue}{Given the pre-critical reading times differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them. Therefore, we refrain from discussing effects occurring in these regions.}}\label{only_precritical} The effects in the pre-critical region are discussed next.

@@

\textcolor{blue}{To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region. This model included fixed effects for semantic interference (high +0.5, low -0.5), trial id (recoded to span from 0 to 1) and the centered plausibility ratings. We used the same priors as presented in Table \ref{tab:spr_priors}. The model was run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items. Figure \ref{fig:posteriors_plausibility} presents the posteriors from the model with the Normal(0, 0.05) prior on the slopes. Although plausibility influenced reading times (higher plausibility lead to faster reading times, CrI [-30, -6] ms), there was an independent effect of semantic interference (CrI [2, 24] ms). The Bayes factors provided evidence for semantic interference in this analysis under all priors (Normal(0, 0.01): BF$_{10}$ = 4.2,
Normal(0, 0.05): BF$_{10}$ = 2.4,
Normal(0, 0.1): BF$_{10}$ = 1.3). The interaction of semantic interference and plausibility was centered around zero. In conclusion, the slowdown in the reading time results for high vs.\ low semantic interference was partially caused by reduced plausibility of the high semantic interference conditions. However, the plausibility difference was not the only driver of the reading time slowdown, i.e., semantic interference leads to slower reading times independent of plausibility.}

@@

The semantic interference effect is further illustrated in Figure \ref{fig:erp_sem_syn} (a) and (b). Figure \ref{fig:erp_sem_syn} (a) shows that around 400 ms, the critical verb elicited a more negative ERP under high than under low semantic interference. Figure \ref{fig:erp_sem_syn} (b) shows that the topography of the effect is rather broad but with a concentration of more negative values at centro-parietal electrodes. The timing and topography of this effect suggested that it was an N400 effect; thus, the N400 was modulated by semantic interference. \textcolor{blue}{Furthermore, Figure \ref{fig:erp_sem_syn} (a) shows a slightly reduced P600 amplitude for high vs.\ low semantic interference. See Figure \ref{fig:erp_sem_syn} (c) for the topographic distribution of this effect.} \Copy{ERP_results2}{In contrast, Figure \ref{fig:erp_sem_syn} (d), (e) and (f) show that syntactic interference did not considerably affect the brain response to the critical verb in \textcolor{blue}{neither the N400 nor P600 spatio-temporal windows}.}\label{ERP_results2} Visual inspection of the ERPs in \ref{fig:erp_sem_syn} (d) does also not suggest that there was a syntactic interference effect in any other time window.

@@

Bayes factors are shown in Figure \ref{fig:eeg_bfs}. \Copy{ERP_results3}{\textcolor{blue}{In the N400 spatio-temporal window, the Bayes factors provided moderate to strong evidence for the semantic interference effect under all priors 
(Normal(0, 0.1): BF$_{10}$ = 6,
Normal(0, 0.5): BF$_{10}$ = 17.5,
Normal(0, 1): BF$_{10}$ = 9.4,
Normal(0, 2): BF$_{10}$ = 5). In the P600 spatio-temporal window, the Bayes factors provided anecdotal to moderate evidence (BF$_{10}$ between 2 and 4.3) for the semantic interference effect under priors assuming effects smaller than $\pm$1\,$\mu V$ and very weak evidence for it (BF$_{10}$ = 1.1) under priors assuming effects in the range of $\pm$2\,$\mu V$. By contrast, Bayes factors provided evidence against the syntactic interference effect in both spatio-temporal windows (BF$_{10}$ $<$ 0.9). The only exception was under the narrowest prior in the N400 window, there was very weak evidence for syntactic interference (BF$_{10}$ = 1.2). Similarly, Bayes factors provided evidence against the interaction in both spatio-temporal windows. The only exception was under the narrowest prior in the P600 window, there was very weak evidence for the interaction (BF$_{10}$ = 1.2). In sum, in our ERP data, there was decisive evidence for the semantic interference effect and mainly evidence against syntactic interference and the interaction.}}\label{ERP_results3}

@@

\Copy{ERP_discussion}{The Bayes factors analyses found \textcolor{blue}{moderate to strong evidence for semantic interference in the N400 amplitude for the whole range of investigated effects sizes ($\pm$0.2 to $\pm4$\,$\mu V$). In the P600 spatio-temporal window, there was anecdotal to moderate evidence for semantic interference. Both the N400 and P600 were more negative under high vs. low semantic interference. In contrast, there was only very weak evidence for very small effects of syntactic interference and the interaction or even evidence against them in both the N400 and P600 windows. In sum, the ERP results showed clear evidence for semantic interference and mostly evidence against syntactic interference and the interaction of syntactic and semantic interference. Consequently, we conclude that the ERP data showed no syntactic interference and no interaction.}}\label{ERP_discussion}

Our finding that retrieval interference modulates the N400 amplitude supports the N400 as a marker of memory retrieval \citep{kutas&federmeier_2000, kutas_federmeier2011, brouwer2017_n4_p6, lau2008_n400}. The more complex memory retrieval under high semantic interference resulted in a more negative N400. \textcolor{blue}{This finding is in line with previous ERP studies on interference \citep{lee_garnsey, vasishth_drenhaus_2011, martinetal2014, schoknecht2022}. The semantic interference effect on the P600 is in the opposite direction than expected: High semantic interference lead to a reduced P600 amplitude compared to low semantic interference. Typically, the P600 amplitude is more positive when syntactic processing is complex, e.g., due to reanalysis \citep{osterhout&holcomb_1992}. The reduced P600 amplitude under high semantic interference in the present study might be explained by facilitatory interference \citep[see e.g.,][]{jaeger_etal_2017}. In the present design, subject-verb integration might be facilitated, i.e., easier, under high semantic interference because there are two semantically suitable candidates to act as the subject. A similar effect was found by \textcite{Tanner_etal_2017}. In their study, the P600 elicited by ungrammatical verbs which were incongruent in number was reduced when the sentence included a matching distractor. However, the present study investigated only grammatical subject-verb integration and the size of the P600 effect was rather small. The small size of the effect on average could indicate that the facilitation occurred only occasionally when the distractor was not ruled out during memory retrieval prior to integration. Since the Bayes factor analysis provided stronger support for the semantic interference effect in the N400 (largest BF$_{10}$ = 17.5) than in the P600 (largest BF$_{10}$ = 4.3), we focus on the N400 results in the remainder of this paper.}

Consistent with the self-paced reading study, there is clear evidence for semantic interference, and \textcolor{blue}{no evidence for syntactic interference in our ERP data}. How do these observed patterns compare to the theoretical predictions of cue-based retrieval theory? We turn to this point next by discussing the quantitative predictions of the \textcite{Lewis2005} model.

@@

\label{modeling}\Copy{modeling}{\textcolor{blue}{The \citet{Lewis2005} model is generally used to predict reading times. Recently, \textcite{mertzen} presented simulations for the present design that predicted approximately equally sized effects of syntactic and semantic interference in the reading times of the critical verb. We found that the reading times at the critical verb were confounded by pre-critical effects, which were best explained by \textit{encoding} interference. Because  Lewis and Vasishth (2005) model is a model of retrieval interference and not encoding interference, and because we cannot disentangle encoding and retrieval interference in the present design, it would not make sense to compare the results from the present SPR experiment to the model's predictions. However, the ERP data can be compared to the model's predictions; all that has to be changed is that the measurement is now on the microvolt scale instead of the millisecond scale. In order to model the ERP data, we therefore adjust the scaling factor $F$, which rescales the activation of items in memory to the relevant dependent measure. The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data. }

>>>>

Semantic interference is manipulated via the animacy of the distractor. During the retrieval of the subject of the verb \textit{trank} (`drank'), the animate distractor  \textit{Einbrecher} (`burglar') is assumed to cause high semantic interference, while the inanimate distractor \textit{Verlust} (`loss') is assumed to cause low semantic interference. \Copy{semantic_features_encoding}{\textcolor{blue}{The animacy of the distractor is used as a stand-in feature for semantic features relevant for the specific retrieval event which might be more fine-grained. In \ref{ex:materials}, the semantic retrieval cue might be \{+animate\} or \{+human\} but it might as well be \{+can drink\}. While more generic features like \{+animate\} or \{+human\} (all animate distractors were human) might be created during memory encoding, more specific features like \{+can drink\} are probably only activated when relevant. The present study is not designed to provide insight into the specific semantic features used during memory encoding and retrieval. For experimental work on highly specific retrieval cues like \{+sailable\} and \{+shatterable\}, see \textcite{vandyke_mcelree06, cunnings_sturt_2018retrieval}. For a principled approach to feature selection using word embeddings, see \textcite{smith2020principled}. A plausibility norming experiment (see page \pageref{plausib_normin_orig}) was conducted to ensure that the semantic manipulation did work, i.e., that the semantic requirements of the verb matched the features of the subject and the animate distractor, but not the features of the inanimate distractor.
}}\label{semantic_features_encoding}

@@

Was there any evidence for the effects \textcolor{blue}{of interest }estimated above? The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}. The analyses use all the data from Experiments 1a and 1b. \Copy{SPR_bfs}{\textcolor{blue}{We found very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231). In the critical region, the evidence for semantic interference was weak (Normal(0, 0.01): BF$_{10}$ = 3.5,
Normal(0, 0.05): BF$_{10}$ = 1.3,
Normal(0, 0.1): BF$_{10}$ = 0.7). There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5). Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8). Similarly,} \Copy{insig_interaction}{\textcolor{blue}{there was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1). Only under the narrowest prior in the critical region, there was very weak evidence for the interaction (BF$_{10}$ = 1.4).}}\label{insig_interaction}

\textcolor{blue}{All in all, the Bayes factor analysis provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but at best weak evidence in the critical region. In contrast, the Bayes factor analysis provided at best weak evidence for syntactic interference and the interaction or even evidence against them.}}

\Copy{fig_bf_spr}{\begin{figure}[htpb]
    \caption{Bayes Factors for the effects of semantic interference, syntactic interference and their interaction in the reading times (combined Experiments 1a and 1b) at the critical verb and surrounding regions, under a range of priors on the target parameters.}
    \label{fig:spr_bfs}
    \centering
    \includegraphics[width=\textwidth]{images/BF_plot_spr_774_allregions.png}
\end{figure}
}

\subsection{Discussion}
We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation. \textcolor{blue}{The reading time results at the critical verb showed anecdotal evidence against or moderate evidence for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region. This means that semantic interference occurred even before the retrieval event could begin. Reading times of the pre-critical word and critical verb showed evidence against syntactic interference. In the spill-over region, there was weak evidence for a small syntactic interference effect in the size of up to $\pm$8 ms. Reading times at the critical verb showed very weak support for a small interaction of syntactic and semantic interference. There was evidence against the interaction under wider priors and in the surrounding regions.} \Copy{only_precritical}{\textcolor{blue}{Given the pre-critical reading times differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them. This means that although it is interesting that small effects of syntactic interference and the interaction emerged only in these regions  any interpretation of these effects would be purely speculative. Therefore, we refrain from discussing the effects occurring in the critical and spill-over regions.}}\label{only_precritical} \textcolor{blue}{We will discuss the absence of decisive evidence for syntactic interference in our data in the general discussion.} The effects in the pre-critical region are discussed next.

@@

\textcolor{blue}{To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region. This model included fixed effects for semantic interference (high +0.5, low -0.5), trial id (recoded to span from 0 to 1) and the centered plausibility ratings. We used the same priors as presented in Table \ref{tab:spr_priors}. The model was run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items. Figure \ref{fig:posteriors_plausibility} presents the posteriors from the model with the Normal(0, 0.05) prior on the slopes. Although plausibility influenced reading times (higher plausibility lead to faster reading times, CrI [-30, -6] ms), there was an independent effect of semantic interference (CrI [2, 24] ms). The Bayes factors provided evidence for semantic interference in this analysis under all priors except the most diffuse one (Normal(0, 0.01): BF$_{10}$ = 4.2,
Normal(0, 0.05): BF$_{10}$ = 2.4,
Normal(0, 0.1): BF$_{10}$ = 1.3). The interaction of semantic interference and plausibility was centered around zero. In conclusion, the slowdown in the reading time results for high vs.\ low semantic interference was partially caused by reduced plausibility of the high semantic interference conditions. However, the plausibility difference was not the only driver of the reading time slowdown, i.e., semantic interference leads to slower reading times independent of plausibility.}

@@

The semantic interference effect is further illustrated in Figure \ref{fig:erp_sem_syn} (a) and (b). Figure \ref{fig:erp_sem_syn} (a) shows that around 400 ms, the critical verb elicited a more negative ERP under high than under low semantic interference. Figure \ref{fig:erp_sem_syn} (b) shows that the topography of the effect is rather broad but with a concentration of more negative values at centro-parietal electrodes. The timing and topography of this effect suggested that it was an N400 effect; thus, the N400 was modulated by semantic interference. \textcolor{blue}{Furthermore, Figure \ref{fig:erp_sem_syn} (a) shows a slightly reduced amplitude for high vs.\ low semantic interference in the P600 time window. Figure \ref{fig:erp_sem_syn} (c) shows that this effect had a broad distribution.} \Copy{ERP_results2}{\textcolor{blue}{Figure \ref{fig:erp_sem_syn} (d) shows that syntactic interference affected the brain response to the critical verb in the N400 and P600 spatio-temporal windows to a smaller extend than semantic interference did. Figures \ref{fig:erp_sem_syn} (d) and (e) show that high vs.\ low syntactic interference lead to a slightly more negative ERP response with a broad distribution in the N400 time window. Figure  \ref{fig:erp_sem_syn} (f) shows a slightly reduced P600 amplitude for high vs.\ low syntactic interference with a centro-parietal distribution. }}\label{ERP_results2}

@@

Bayes factors are shown in Figure \ref{fig:eeg_bfs}. \Copy{ERP_results3}{\textcolor{blue}{In the N400 spatio-temporal window, the Bayes factors provided moderate to strong evidence for the semantic interference effect under all priors 
(Normal(0, 0.1): BF$_{10}$ = 6,
Normal(0, 0.5): BF$_{10}$ = 17.5,
Normal(0, 1): BF$_{10}$ = 9.4,
Normal(0, 2): BF$_{10}$ = 5). In the P600 spatio-temporal window, the Bayes factors provided anecdotal to moderate evidence (BF$_{10}$ between 2 and 4.3) for the semantic interference effect under priors assuming effects smaller than $\pm$1\,$\mu V$ and no evidence for it (BF$_{10}$ = 1.1) under priors assuming effects in the range of $\pm$2\,$\mu V$. By contrast, Bayes factors provided either no evidence for or even evidence against syntactic interference in both spatio-temporal windows (BF$_{10}$ $<$ 1.2). Similarly, Bayes factors provided either no evidence for or even evidence against the interaction in both spatio-temporal windows (BF$_{10}$ $<$ 1.2). In sum, in our ERP data, there was decisive evidence for the semantic interference effect and mainly evidence against syntactic interference and the interaction.}}\label{ERP_results3}

@@

\Copy{ERP_discussion}{The Bayes factors analyses found \textcolor{blue}{moderate to strong evidence for semantic interference in the N400 amplitude for the whole range of investigated effects sizes ($\pm$0.2 to $\pm4$\,$\mu V$). In the P600 spatio-temporal window, there was anecdotal to moderate evidence for semantic interference. Both the N400 and P600 were more negative under high vs. low semantic interference. In contrast, there was no evidence for effects of syntactic interference and the interaction or even evidence against them in both the N400 and P600 windows. In sum, the ERP results showed clear evidence for semantic interference and mostly evidence against syntactic interference and the interaction of syntactic and semantic interference. Consequently, we conclude that the ERP data showed no syntactic interference and no interaction.}}\label{ERP_discussion}

Our finding that retrieval interference modulates the N400 amplitude supports the N400 as a marker of memory retrieval \citep{kutas&federmeier_2000, kutas_federmeier2011, brouwer2017_n4_p6, lau2008_n400}. The more complex memory retrieval under high semantic interference resulted in a more negative N400. \textcolor{blue}{This finding is in line with previous ERP studies on interference \citep{lee_garnsey, vasishth_drenhaus_2011, martinetal2014, schoknecht2022}. The semantic interference effect on the P600 is in the opposite direction than expected: High semantic interference lead to a reduced P600 amplitude compared to low semantic interference. Typically, the P600 amplitude is more positive when syntactic processing is complex, e.g., due to reanalysis \citep{osterhout&holcomb_1992}. The reduced P600 amplitude under high semantic interference in the present study might be explained by facilitatory interference \citep[see e.g.,][]{jaeger_etal_2017}. In the present design, subject-verb integration might be facilitated, i.e., easier, under high semantic interference because there are two semantically suitable candidates to function as the subject. A similar effect was found by \textcite{Tanner_etal_2017}. In their study, the P600 elicited by ungrammatical verbs which were incongruent in number was reduced when the sentence included a matching distractor. However, the present study investigated only grammatical subject-verb integration and the size of the P600 effect was rather small. The small size of the effect on average could indicate that the facilitation occurred only occasionally when the distractor was not ruled out during memory retrieval prior to integration. Since the Bayes factor analysis provided stronger support for the semantic interference effect in the N400 (largest BF$_{10}$ = 17.5) than in the P600 (largest BF$_{10}$ = 4.3), we focus on the N400 results in the remainder of this paper.}

\textcolor{blue}{Similar to the self-paced reading results,} there is clear evidence for semantic interference, and \textcolor{blue}{no evidence for syntactic interference in our ERP data}. How do these observed patterns compare to the theoretical predictions of cue-based retrieval theory? We turn to this point next by discussing the quantitative predictions of the \textcite{Lewis2005} model.

@@

\label{modeling}\Copy{modeling}{\textcolor{blue}{The \citet{Lewis2005} model is generally used to predict reading times. Recently, \textcite{mertzen} presented simulations for the present design that predicted approximately equally sized effects of syntactic and semantic interference in the reading times of the critical verb. We found that the reading times at the critical verb were confounded by pre-critical effects, which were best explained by \textit{encoding} interference \citep[see also][]{mertzen}. Because  Lewis and Vasishth (2005) model is a model of retrieval interference and not encoding interference, and because we cannot disentangle encoding and retrieval interference in the present design, it would not make sense to compare the results from the present SPR experiment to the model's predictions. However, the ERP data can be compared to the model's predictions; all that has to be changed is that the measurement is now on the microvolt scale instead of the millisecond scale. In order to model the ERP data, we therefore adjust the scaling factor $F$, which rescales the activation of items in memory to the relevant dependent measure. The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.}

We sampled the free scaling parameter $F$ from a \textcolor{blue}{truncated normal prior distribution ($Normal_{lb=0,ub=0.05}(0.01, 0.01)$) with a lower bound of 0 and an upper bound of 0.05. This prior reflects the a priori belief that effects on the microvolt scale will range from 0 to 5 $\mu V$ with higher likelihood for effect sizes below 3 $\mu V$ (see Figure \ref{fig:model_predictions_ERP}). This relatively uninformative prior was chosen because, although there exists no previously published ERP data on the Van Dyke (2007) design, it reflects the reasonable range of effect sizes in psycholinguistic ERP research.} After sampling the scaling parameter, simulated data was generated from the model while holding all other parameters in the model at fixed values. \textcolor{blue}{We decided to model N400 amplitude differences because Bayes factors provided stronger evidence for the semantic interference effect on the N400 than on the P600 in the present ERP study. Since the scaling parameter $F$ cannot be negative, an additional step was needed to reflect that larger predicted effect sizes correspond to more negative N400 amplitudes: The predicted effects from the model} were multiplied with $-1$. \textcolor{blue}{In future work, the posterior distribution of the semantic interference effect in our ERP data, with range [-0.5, -0.1]\,$\mu V$, could be used as a basis for redefining a tighter prior on the scaling parameter $F$.}
